---
title: "Intake Form Analysis"
author: "Hume Dickie"
date: "December 10, 2017"
output: html_document
---

```{r libs, include=FALSE}
require(googlesheets)
require(magrittr)
require(ggplot2)
require(gridExtra)
require(stringdist)
require(stringr)
require(gplots)
```

```{r setup, include=FALSE}
intake.form.sheetID <- '1eUxxEVagFv4nTGI4MMCFWXUiuIT2deUCs6-q__r19ko'
first.valid.date <- strptime("2016-08-26",format="%Y-%m-%d") %>% as.POSIXct() %>% as.integer()

gs_auth(token = "googlesheets_token.rds")
gs.data.object <- gs_key(intake.form.sheetID)
intake.data <- gs.data.object %>% gs_read(ws=1)

bar.color <- "#fdbb30"
text.color <- "#00543c"
```

```{r configVariables, include=FALSE}
max.appt.count <- 600
repeat.assignment.cutoff <- 5
```

### Appointment Count Over Time For Each Semester
TODO save individual semester plots in a subfolder
```{r betterTimeSeriesPlot, echo=FALSE}
dates  <- intake.data$Date %>% strptime("%m/%d/%Y") %>% as.character()
dates <- dates[which(dates > as.Date("2016-01-01"))]

daily.counts <- table(dates %>% as.character())
semester.days <- seq(as.Date("2016-8-1"), as.Date("2017-12-31"), by="days") %>% as.character()
sel <- semester.days %in% (dates %>% as.character())

for (zeroDay in semester.days[!sel]) {
  daily.counts[zeroDay] <- 0
}

#sortsel
dates <- daily.counts %>% names()
values <- daily.counts %>% as.integer()

daily.time.series <- cbind(dates,values) %>% as.data.frame()
daily.time.series[,1] <- daily.time.series[,1] %>% as.Date()
daily.time.series$values <- daily.time.series$values %>% as.character %>% as.numeric()

daily.time.series <- daily.time.series[order(daily.time.series$dates),]
daily.time.series$dateAsInt <- daily.time.series[,"dates"] %>% as.integer()


#cut into semesters
spring.starts <- seq(as.Date("2013/01/01"), Sys.Date(), by="years")
spring.ends <- seq(as.Date("2013/05/31"), Sys.Date(), by="years")
fall.starts <- seq(as.Date("2013/08/01"), Sys.Date(), by="years")
fall.ends <- seq(as.Date("2013/12/31"), Sys.Date() + 180, by="years")

define.semesters <- c(spring.starts,spring.ends,fall.starts,fall.ends) %>% sort()

daily.time.series$sem <- rep("",nrow(daily.time.series))
for (i in 1:(length(define.semesters)/2)) {
  semester.data <- daily.time.series[which(as.Date(daily.time.series$dates) > define.semesters[[2*i-1]] & as.Date(daily.time.series$dates) < define.semesters[[2*i]]),]

  if(nrow(semester.data) == 0) {
    next
  }
  
  if (i %% 2 == 0) {
    semesterString <- paste("Spring ",semester.data[1,"dates"] %>% substr(1,4),sep="")
  } else {
    semesterString <- paste("Fall ",semester.data[1,"dates"] %>% substr(1,4),sep="")
  }
  
  daily.time.series[which(as.Date(daily.time.series$dates) > define.semesters[[2*i-1]] & as.Date(daily.time.series$dates) < define.semesters[[2*i]]),"sem"] <- semesterString
  
}

daily.time.series <- daily.time.series[which(daily.time.series$sem != ""),]
sem.count.table <- table(daily.time.series$sem)
non.empty.semesters <- names(sem.count.table)[sem.count.table > 10]

#date to int and cumsum
for(s in non.empty.semesters){
    daily.time.series[which(daily.time.series$sem==s),"dateAsInt"] <- daily.time.series[which(daily.time.series$sem==s),"dateAsInt"] - min(daily.time.series[which(daily.time.series$sem==s),"dateAsInt"])
    daily.time.series[which(daily.time.series$sem==s),"cumsum"] <- daily.time.series[which(daily.time.series$sem==s),"values"] %>% cumsum()
}


semester.time.series.plots <- list()
overallPlot <- NULL
for (i in 1:length(non.empty.semesters)) {
  semester.time.series.plots[[i]] <- ggplot(aes(x=dateAsInt,y=cumsum,color=sem),data=daily.time.series[which(daily.time.series$sem==non.empty.semesters[[i]]),]) + geom_line(size=1.5)
  
  date.index.0 <- daily.time.series[which(daily.time.series$sem == non.empty.semesters[[i]] & daily.time.series$dateAsInt == 0),"dates"]
  
  x.label <- paste("Days Since: ",date.index.0,sep="")
  plot.title <- non.empty.semesters[[i]]
  
  semester.time.series.plots[[i]] <- semester.time.series.plots[[i]] + labs(title=plot.title,subtitle=paste(daily.time.series[which(daily.time.series$sem==non.empty.semesters[[i]]),]$cumsum %>% max()," total appointments",sep="")) + labs(x=x.label) + labs(y="Appt Count")
  
}

overallPlot <- ggplot(aes(x=dateAsInt,y=cumsum,color=sem),data=daily.time.series) + geom_line(size=1.5) +xlim(20,135) + ylim(0,max.appt.count) + labs(x="Days from January 2 or August 2",y="Appt Count") + ggtitle("Appointments Over Time",subtitle=paste("Total Appointments: ",sum(daily.time.series$values),sep=""))

for(plot in semester.time.series.plots) {
  print(plot)
}

print(overallPlot)
```

### Types of Assignments

```{r cleanStrings, echo=FALSE}
unique.assignments <- intake.data$Assignment %>% as.character() %>% tolower()  %>% unique()

#at beginning or with whitespace on both sides
filler.words <- list(c("of","in","a","to","the","an","or","and","on","about"))
for (pat in filler.words %>% unlist()) {
  pat <- paste(" ",pat," ",sep="") #in middle of sentence
  unique.assignments <- gsub(pat," ",x=unique.assignments,ignore.case=TRUE)
  
  pat <- trimws(pat)
  pat <- paste("^",pat," ",sep="") #start of line
  unique.assignments <- gsub(pat,"",x=unique.assignments,ignore.case=TRUE)
  
  pat <- trimws(pat)
  pat <- paste(" ",pat,"$",sep="") #end of line
  unique.assignments <- gsub(pat,"",x=unique.assignments,ignore.case=TRUE)
}

#special case
unique.assignments <- gsub("/"," ",x=unique.assignments,ignore.case=TRUE)

#irrelevant words and symbols
delete.patterns <- list(c("speech","presentation","project","[0-9]","#","[[:punct:]]","^n.*a$"))
for (pat in delete.patterns %>% unlist()) {
  unique.assignments <- gsub(pat,"",x=unique.assignments,ignore.case=TRUE)
}

unique.assignments <- gsub("[ ]+"," ",unique.assignments)
unique.assignments <- trimws(unique.assignments)

#some need to be toward the end
unique.assignments <- unique.assignments[unique.assignments !=""]
unique.assignments <- unique.assignments[!is.na(unique.assignments)]
unique.assignments <- unique.assignments[nchar(unique.assignments %>% unlist()) > 2]

unique.assignments <- unique.assignments %>% unique()
```

```{r getLevenshteinDistanceMatrix, echo=FALSE}
getStringDistMatrix <- function(list.of.strings){
  unique.strings <- unique(list.of.strings %>% unlist())
  
  n <- length(unique.strings)
  string.dist.matrix <- matrix(nrow=n,ncol=n)
  
  start <- Sys.time()
  for (i in 1:n) {
    for (j in i:n) {
      dist <- stringdist(unique.strings[[i]],unique.strings[[j]])
      string.dist.matrix[i,j] <- dist
      string.dist.matrix[j,i] <- dist
      
      if (n*i/2 + j %% 24600 == 0) {
        partial <- Sys.time()
        print((n*i + j)/(n**2))
      }
    }
  }
  
  string.dist.matrix[is.na(string.dist.matrix)] <- 0
  string.dist.matrix <- string.dist.matrix %>% apply(2,as.integer)
  return(string.dist.matrix)
}
```

```{r getDistFromSet}
#instead of the distance from each other, use this method to get the distance from strings you know ahead of time should appear
getDistFromSetMatrix <- function(list.of.strings,known.strings) {
  unique.strings <- unique(list.of.strings %>% unlist())
  
  #defines an m x n matrix, where each column represents a known string
  m <- length(unique.strings)
  n <- length(known.strings)
  string.dist.matrix <- matrix(nrow=m,ncol=n)
  
  #for each row, compute dist from each known string
  for (i in 1:m) {
    for (j in 1:n) {
      dist <- stringdist(unique.strings[[i]],known.strings[[j]])
      string.dist.matrix[i,j] <- dist
    }
  }
  
  
  string.dist.matrix[is.na(string.dist.matrix)] <- 0
  string.dist.matrix <- string.dist.matrix %>% apply(2,as.integer)
  return(string.dist.matrix)
}

```

```{r clusterStrings, echo = FALSE}
full.dist.matrix <- getStringDistMatrix(unique.assignments)
hc.assignments <- hclust(as.dist(full.dist.matrix))
```

```{r stringDistInsights, echo=FALSE}
sizeOfGroup1 <- list()
diff.size <- list()
for (cutk in 1:length(unique.assignments)) {
  sizeOfGroup1[[cutk]] <- which(cutree(hc.assignments,k=cutk) == 1) %>% length()
  
  if(cutk>1) {
    diff.size[[cutk-1]] <- sizeOfGroup1[[cutk]] - sizeOfGroup1[[cutk-1]]
  }
}

optimal.k <- which(diff.size == diff.size %>% unlist() %>% min()) + 1


```

```{r computeFinalOutput, include=FALSE}
repeat.assignment.cutoff <- 5

clusters.assignments <- data.frame(unique.assignments,cutree(hc.assignments,k=optimal.k))
assignment.table <- clusters.assignments$cutree.hc.assignments..k...optimal.k. %>% table()
repeat.assignments <- assignment.table[assignment.table > repeat.assignment.cutoff]
repeat.assignments <- repeat.assignments[repeat.assignments != max(repeat.assignments)]
#the largest one is usually nonsense
repeat.indices <- repeat.assignments %>% sort() %>% rev() %>% names() %>% as.numeric()

```


This is common types of assignments, but not how many people come in with assignments of this type
```{r outputClusters, echo=FALSE}
for (i in 1:length(repeat.indices)) {
  relevant.assignments.df <- clusters.assignments[which(clusters.assignments$cutree.hc.assignments..k...optimal.k. == repeat.indices[[i]]),]
  
  print("-----------------------------------------------------")
  relevant.assignments.df$unique.assignments %>%as.character() %>% print()
}
```


Now, let's make a bar chart of specific assignments that we're looking out for.
I think I should grep to find composites of atomic assignment descriptors
```{r plotKnownAssignments}
known.assignments <- c("persuasive","informative","rogerian","paper","essay","final","extemporaneous","outline","comemmorative","culture","definition","first","group","introduction","memorize","problem solution","proposal","research","special occasion","storytelling","life story","praise blame","toefl","significant life story","research paper","significant life","epideictic","memorization","essay outline","final draft","oral interpretation","debate","final paper")

dist.from.known.assignments <- getDistFromSetMatrix(unique.assignments,known.assignments)

pretty.close <- which(dist.from.known.assignments <= 4)
near.match <- (pretty.close/length(unique.assignments)) %>% ceiling()
input.assignment  <- pretty.close %% length(unique.assignments)

unique.assignments[input.assignment]

unmatched <- rep(TRUE,length(unique.assignments))
unmatched[input.assignment] <- FALSE

unique.assignments[unmatched]
```

### Professors

```{r cleanStringsProfs, echo=FALSE}
unique.profs <- intake.data$Professor %>% as.character() %>% tolower()  %>% unique()

delete.patterns <- list(c("professor","prof","[0-9]","#","[[:punct:]]","^n.*a$"))
for (pat in delete.patterns %>% unlist()) {
  unique.profs <- gsub(pat,"",x=unique.profs,ignore.case=TRUE)
}

unique.profs <- gsub("[ ]+"," ",unique.profs)
unique.profs <- trimws(unique.profs)

#some need to be toward the end
unique.profs <- unique.profs[unique.profs !=""]
unique.profs <- unique.profs[!is.na(unique.profs)]
unique.profs <- unique.profs[nchar(unique.profs %>% unlist()) > 2]

unique.profs <- unique.profs %>% unique()
```

Let's do substring matches searching for last names only!
String distance from split on space! If one of them is super low call it a match!

Angela Beasley
Beste Yuksel
Cynthia Shultes
Deborah Callister
Eric Christiansen
Gina Stahlricco
Jacquelyn Horton
Leigh Meredith
Lindsey Ayotte
Lynn Perkins
Maximilian Delaure
Michael Rozendal
Patricia Kalman
Suzanne Pullens
```{r distFromSetsProfs}
#usf.professors <- c("hunt","horton","delore","delaure","beasly","callister","christian","ricco","rosendahl","ayotte","perkins","pullen","beasley","martin","meredith","sullivan","rozendal","beasely","yuksel","rosendal","stahl","kalman","gina","stahlricco","schultes","suzanne","cynthia")

usf.professors <- c("angela","beasley","beste","yuksel","cynthia","shultes","deborah","callister","eric","christiansen","gina","stahlricco","jacquelyn","horton","leigh","meredith","lindsey","ayotte","lynn","perkins","maximilian","delaure","michael","rozendal","patricia","kalman","suzanne","pullens")

unmatched.profs <- rep(TRUE,length(unique.profs))
mapped.profs <- rep("",length(unique.profs))
for(i in 1:length(unique.profs)) {
  for(professor in usf.professors) {
    #print(paste(professor," in ",unique.profs[[i]],"?",sep=""))
    if (grepl(professor,unique.profs[[i]])) {
      unmatched.profs[[i]] <- FALSE
      mapped.profs[[i]] <- professor
    }
  }
}

#unmatched and has a space in it
unique.profs[unmatched.profs & (lapply(unique.profs,grepl,pattern=" ") %>% unlist())] %>% sort()

#same length as unique.rpfs, but every element is here a list of one or more strings
for(i in 1:length(unique.profs)) {
  splitname <- strsplit(unique.profs[[i]]," ")
  for (partialname in splitname) {
    if(!unmatched.profs[[i]]){
      next
    }
    
    dist.from.each.prof <- getDistFromSetMatrix(usf.professors,partialname)
    
    near.matches.indices <- which(dist.from.each.prof <= 3) %% length(usf.professors) %>% unique()
    mapped.profs[[i]] <- usf.professors[near.matches.indices] %>% list()
    
    if(length(mapped.profs) > 0) {
      unmatched.profs[[i]] <- FALSE
    }
  }
}

unique.profs[unmatched.profs]
```

```{r clusterStringsProfs, echo = FALSE}
full.dist.matrix.profs <- getStringDistMatrix(unique.profs)
hc.profs <- hclust(as.dist(full.dist.matrix.profs))
```

```{r stringDistInsightsProfs, echo=FALSE}
sizeOfGroup1 <- list()
diff.size <- list()
for (cutk in 1:length(unique.profs)) {
  sizeOfGroup1[[cutk]] <- which(cutree(hc.profs,k=cutk) == 1) %>% length()
  
  if(cutk>1) {
    diff.size[[cutk-1]] <- sizeOfGroup1[[cutk]] - sizeOfGroup1[[cutk-1]]
  }
}

optimal.k <- which(diff.size == diff.size %>% unlist() %>% min()) + 1


```

```{r computeFinalOutputProfs, include=FALSE}

clusters.profs <- data.frame(unique.profs,cutree(hc.profs,k=optimal.k))
profs.table <- clusters.profs$cutree.hc.profs..k...optimal.k. %>% table()
repeat.profs <- profs.table[profs.table > 2]
repeat.profs <- repeat.profs[repeat.profs != max(repeat.profs)]
#the largest one is usually nonsense
repeat.indices <- repeat.profs %>% sort() %>% rev() %>% names() %>% as.numeric()

```


This is common types of assignments, but not how many people come in with assignments of this type
```{r outputClusters{Profs}, echo=FALSE}
for (i in 1:length(repeat.indices)) {
  relevant.profs.df <- clusters.profs[which(clusters.profs$cutree.hc.profs..k...optimal.k. == repeat.indices[[i]]),]
  
  print("-----------------------------------------------------")
  relevant.profs.df$unique.profs %>% as.character() %>% print()
}
```

### Classes

```{r cleanStringsClasses, echo=FALSE}
unique.classes <- intake.data$Class %>% as.character() %>% tolower()  %>% unique()

#at beginning or with whitespace on both sides
#filler.words <- list(c("of","in","a","to","the","an","or","and","on"))
#for (pat in filler.words %>% unlist()) {
#  pat <- paste(" ",pat," ",sep="") #in middle of sentence
#  unique.assignments <- gsub(pat," ",x=unique.assignments,ignore.case=TRUE)
#  
#  pat <- trimws(pat)
#  pat <- paste("^",pat," ",sep="") #start of line
#  unique.assignments <- gsub(pat,"",x=unique.assignments,ignore.case=TRUE)
#  
#  pat <- trimws(pat)
#  pat <- paste(" ",pat,"$",sep="") #end of line
#  unique.assignments <- gsub(pat,"",x=unique.assignments,ignore.case=TRUE)
#}

#special case
unique.classes <- gsub("/"," ",x=unique.classes,ignore.case=TRUE)

##irrelevant words and symbols
#delete.patterns <- list(c("speech","presentation","project","[0-9]","#","[[:punct:]]","^n.*a$"))
#for (pat in delete.patterns %>% unlist()) {
#  unique.assignments <- gsub(pat,"",x=unique.assignments,ignore.case=TRUE)
#}

unique.classes <- gsub("[ ]+"," ",unique.classes)
unique.classes <- trimws(unique.classes)

#some need to be toward the end
unique.classes <- unique.classes[unique.classes !=""]
unique.classes <- unique.classes[!is.na(unique.classes)]
unique.classes <- unique.classes[nchar(unique.classes %>% unlist()) > 2]

unique.classes <- unique.classes %>% unique()
```


```{r clusterStringsClasses, echo = FALSE}
full.dist.matrix.class <- getStringDistMatrix(unique.classes)
hc.classes <- hclust(as.dist(full.dist.matrix.class))
```

```{r stringDistInsightsClasses, echo=FALSE}
sizeOfGroup1 <- list()
diff.size <- list()
for (cutk in 1:length(hc.classes)) {
  sizeOfGroup1[[cutk]] <- which(cutree(hc.classes,k=cutk) == 1) %>% length()
  
  if(cutk>1) {
    diff.size[[cutk-1]] <- sizeOfGroup1[[cutk]] - sizeOfGroup1[[cutk-1]]
  }
}

optimal.k <- which(diff.size == diff.size %>% unlist() %>% min()) + 1


```

```{r computeFinalOutputClasses, include=FALSE}
clusters.classes <- data.frame(unique.classes,cutree(hc.classes,k=optimal.k))
classes.table <- clusters.classes$cutree.hc.classes..k...optimal.k. %>% table()
repeat.classes <- classes.table[classes.table > 1]
repeat.classes <- repeat.classes[repeat.classes != max(repeat.classes)]
#the largest one is usually nonsense
repeat.indices <- repeat.classes %>% sort() %>% rev() %>% names() %>% as.numeric()

```


This is common types of classes, but not how many people come in with assignments of this type
```{r outputClustersClasses, echo=FALSE}
for (i in 1:length(repeat.indices)) {
  relevant.classes.df <- clusters.classes[which(clusters.classes$cutree.hc.classes..k...optimal.k. == repeat.indices[[i]]),]
  
  print("-----------------------------------------------------")
  relevant.classes.df$unique.classes %>%as.character() %>% print()
}
```

### Sign Up Method & Extra Credit/Required
  TODO over time     
      
```{r summaryStats, echo=FALSE}
reservedThrough <- table(intake.data$`Appointment Reserved Through`)
required <- table(intake.data$`Is this visit required by your professor?`)
#unique visitors
#percent capacity

reservedThrough.df <- cbind(reservedThrough %>% names(), reservedThrough %>% as.vector()) %>% as.data.frame()
required.df  <- cbind(required %>% names(), required %>% as.vector()) %>% as.data.frame()

reservedThrough.df$V2 <- reservedThrough.df$V2 %>% as.character() %>% as.numeric()
required.df$V2 <- required.df$V2 %>% as.character() %>% as.numeric()

x.label <- (reservedThrough.df$V2/sum(reservedThrough.df$V2)) %>% as.character() %>% substr(3,4) %>% paste("%",sep="")

ggplot(reservedThrough.df,aes(x=V1,y=V2)) + geom_bar(stat="identity",fill=bar.color) + labs(x="Appointment Sign Up Method") + labs(y="Percent of Total")  + 
    geom_text(data=reservedThrough.df,aes(x=V1,y=V2 %>% as.numeric(),label=x.label %>% as.character()),col=text.color,vjust=-0.25)

x.label <- (required.df$V2/sum(required.df$V2)) %>% as.character() %>% substr(3,4) %>% paste("%",sep="")

ggplot(required.df,aes(x=V1,y=V2)) + geom_bar(stat="identity",fill=bar.color) + labs(x="Requirement/Extra Credit Status") + labs(y="Percent of Total")  + 
    geom_text(data=required.df,aes(x=V1,y=V2 %>% as.numeric(),label=x.label %>% as.character()),col=text.color,vjust=-0.25)
```

### Unique and Returning Students

### Classes outside of Rhet

### Reasons they come in